---
title: "Toronto Fire Services Incidents Revisitied"
author: "Geoffrey Clark"
date: "10/2/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Data
A lot of preparation has already been conducted so we'll just load that data here...
```{r}
load("dataprep.RData")

# Set a seed for reproducability:
set.seed(42)
```



## Class Imbalance
There's a pretty big class imbalance here: ~96/4
```{r}
table(I$CRITICAL) / nrow(I) * 100

# Let's first create a data.frame containing features of interest, Incidents subset:
I_s <- I[,c('EVENT_TYPE_CD','ALARM_TO_FD','RESPONSE_TYPE','EST_KM','INITIAL_UNIT_PERSONNEL','INCIDENT_DAY','INCIDENT_MONTH','INITIAL_CALL_HOUR','INITIAL_CALL_MIN','CRITICAL')]
I_s$EVENT_TYPE_CD <- as.factor(I_s$EVENT_TYPE_CD)

# With less features it's easy to oversample the imbalanced class with ROSE (Random Over-sampling Examples).
# For simplicity let's use a previously randomly selected training set (training/validation/test = 60/20/20)

table(I_s[full_idx$train, 'CRITICAL'])/length(full_idx$train)

library(ROSE)

I_s_ROSE <- ROSE(CRITICAL ~ ., data=I_s[full_idx$train,])$data

# Our data now has about a 50/50 class split. Not bad!

table(I_s_ROSE$CRITICAL) / nrow(I_s_ROSE) * 100

# Let's train a model with this new set and see how it performs.

model_logistic_bal <- glm(CRITICAL ~ ., data=I_s_ROSE, family=binomial(link="logit"))
model_logistic_imb <- glm(CRITICAL ~., data = I_s[full_idx$train,], family=binomial(link="logit"))
source("custom_summaries.R")

model_log_bal_predict <- predict.logistic(model_logistic_bal, I_s[full_idx$cv,], prob=0.5)
model_log_imb_predict <- predict.logistic(model_logistic_imb, I_s[full_idx$cv,], prob=0.5)

# source("refold.models.R")
# metric.lr <- refold.models(I_s_ROSE)


```
### Input Data: Responding Units
```{r}
## Responding Units
# Load & Combine Dataset from Yearwise .csv

RU <- read.csv("../csv/2011_ru.csv", header=T, stringsAsFactors = F, na.strings=c("","NA"))


for(yr in (2012:2016)){
  r_file <- paste("../csv/",yr, "_ru.csv", sep="")
  
  r_df <- read.csv(r_file, header=T, stringsAsFactors = F)
  RU <- rbind(RU, r_df)
  rm(r_df)
  rm(r_file)
    
}


```

### Responding Unit Features

The goal here is to retrieve information from the responding units XML
According to contact at Toronto Fire, the first unit dispatched is usually from the station that received the dispatch.


```{r}
RS <- data.frame(INCIDENT_NUMBER=character(),
                 FIRST_STATION=character(),
                 SECOND_STATION=character(),
                 THIRD_STATION=character(),
                 FIRST_APPARATUS=character(),
                 SECOND_APPARATUS=character(),
                 THIRD_APPARATUS=character())


sum(is.na(RU$DISPATCH_TIME)) #46. 
# Small number of observations without DISPATHC_TIME. Removing them for now 

RU <- RU[!is.na(RU$DISPATCH_TIME),]
RU <- RU[grepl("\\d{4}-\\d{2}-\\d{2} \\d{2}:\\d{2}:\\d{2}", RU$DISPATCH_TIME),]


RU$DISPATCH_TIME <- as.POSIXct(RU$DISPATCH_TIME)

curr_year = 0
for(i in I$INCIDENT_NUMBER[nchar(I$INCIDENT_NUMBER) == 9]){
  if(curr_year != substr(i,2,3)){
    curr_year <- substr(i, 2,3)
    cat("\n", curr_year) 
  }
  
  subset <- RU[RU$INCIDENT_NUMBER == i,]
  subset <- subset[order(subset[,3]),]
  
  this_RS <- data.frame(INCIDENT_NUMBER=i,
                 FIRST_STATION=as.character(NA),
                 SECOND_STATION=as.character(NA),
                 THIRD_STATION=as.character(NA),
                 FIRST_APPARATUS=as.character(NA),
                 SECOND_APPARATUS=as.character(NA),
                 THIRD_APPARATUS=as.character(NA))
  
  this_RS$FIRST_STATION <- strsplit(subset[1,'CAD_UNIT_ID'], "\\D+")[[1]][2]
  this_RS$FIRST_APPARATUS <- strsplit(subset[1,'CAD_UNIT_ID'], "\\d+")[[1]][1]
  
  if(nrow(subset) > 1){
    this_RS$SECOND_STATION <- strsplit(subset[2,'CAD_UNIT_ID'], "\\D+")[[1]][2]
    this_RS$SECOND_APPARATUS <- strsplit(subset[2,'CAD_UNIT_ID'], "\\d+")[[1]][1]
  }
  if(nrow(subset) > 2){
    this_RS$THIRD_STATION <- strsplit(subset[3,'CAD_UNIT_ID'], "\\D+")[[1]][2]
    this_RS$THIRD_APPARATUS <- strsplit(subset[3,'CAD_UNIT_ID'], "\\d+")[[1]][1]
    
  }
  
  RS <- rbind(RS, this_RS)
  
  
}
```

